---
title: "Slurm Job Assessment"
output:
  html_document:
    code_folding: hide
  pdf_document: default
date: "`r Sys.Date()`"
params:
  user: NA
  account: cedar,cedar2
  start: NA
  end: NA
  file: NA
  all: 'TRUE'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r message=FALSE}
library(data.table, quietly = TRUE)
library(stringr)
library(tidyverse)
library(knitr)
library(DT)
```

## Retrieve Data 
First, checks if file exists. If file doesn't exist, sacct command will be ran. 


```{r}
unlockBinding("params", env = .GlobalEnv)

params <- lapply(params, function(x) { if (x == 'NA' | x == '') NA else x })

file <- params$file 
user <- params$user 
start <- params$start 
end <- params$end
account <- params$account
all <- params$all
output_file <- paste0("clean", file)

if (!is.na(file)){
  print("Using file provided..")
  lines <- readLines(file)
  writeLines(lines[nzchar(trimws(lines))], output_file)
  data <- fread(output_file)
} else { 
  print("Pulling sacct data..")
  # if end is NA -> use today's date 
  if (is.na(end)){ 
     end <- Sys.Date()}   
  # if start is also NA -> use past week from today 
  
  if (is.na(start)){
    start <- as.character(as.Date(end) - 7 ) }
  

  
  sacct_command <- sprintf("sacct --units=G --format=JobIdRaw,JobName,User,Group,Account,State,Submit,Start,End,Cluster,Partition,AllocNodes,AllocTRES,AllocCPUS,ReqCPUs,AveCPU,TotalCPU,CPUTime,UserCPU,SystemCPU,Elapsed,Timelimit,ReqMem,MaxRSS,MaxVMSize,MaxDiskWrite,MaxDiskRead,CPUTimeRaw,ElapsedRaw,TimelimitRaw --parsable2 -a -A %s --starttime=%s --endtime=%s", account, start, end)

  
   if (!is.na(user)){ 
    sacct_command <- paste(sacct_command, sprintf("-u %s", user))
  }
  
  temp_file <- tempfile()
  system(paste(sacct_command, ">", temp_file), wait = TRUE)
  data <- fread(temp_file)
  
  print(sacct_command)
  

}

print(sprintf("File: %s, User: %s, Account: %s, Start: %s, End: %s, all: %s", file, user, account, start, end, all)) 

```
```{r}
if (nrow(data) == 0){ 
  print("Data retrieved is empty! Review sacct command or your inputted file. ")
  knit_exit()
  }
```

```{r}
# Calculate time in seconds - brute force function
calcSeconds <- function(slurmTime) {
  fun.dt <- data.table(dtime = slurmTime)
  mytimecols <- c("dd", "hh", "mm", "ss")
  # Split out days, hours, minutes, and seconds
  try(fun.dt[grepl("-", dtime),
             c("dd", "hh", "mm", "ss") := tstrsplit(dtime, "[-:]")], silent = TRUE)
  try(fun.dt[!grepl("-", dtime) & str_count(dtime, ":") == 2,
             c("dd", "hh", "mm", "ss") := c(0, tstrsplit(dtime, "[-:]"))], silent = TRUE)
  try(fun.dt[!grepl("-", dtime) & str_count(dtime, ":") == 1,
             c("dd", "hh", "mm", "ss") := c(0, 0, tstrsplit(dtime, "[-:]"))], silent = TRUE)
  
  # Convert from chr to num as available
  myfuntime <- mytimecols[mytimecols %in% colnames(fun.dt)]
  fun.dt[, (myfuntime) := lapply(.SD, as.numeric), .SDcols = myfuntime]
  
  # Calculate total time in seconds
  fun.dt[, dtimeSec := dd*24*60*60 + hh*60*60 + mm*60 + ss]
  return(fun.dt$dtimeSec)
}
```



```{r warnings = FALSE}
# Split out jobs & subjobs for convenience
data[, c("JobID", "SubJob") := tstrsplit(JobIDRaw, "\\.")]

# Convert memory to numeric
units <- "G"
for (mycol in c("ReqMem", "MaxRSS", "MaxVMSize", "MaxDiskRead", "MaxDiskWrite")) {
  newcol <- paste(mycol, units, sep = "")
  data[!(get(mycol) %in% c("", "0")), (newcol) := as.numeric(gsub(units, "", get(mycol)))]
}

# Converting TotalCPU and CPUTime to seconds for efficiency calcs 
data[, TotalCPUSec := calcSeconds(TotalCPU)]
data[, CPUTimeSec := calcSeconds(CPUTime)]

```


```{r}

colnames.both <- c("Submit")
colnames.job <- c("JobName", "User", "State", "Group", "Account", "Submit", "Start", "End", "Cluster",
                  "Partition", "AllocTRES", "ReqCPUS", "TotalCPU", "CPUTime", "UserCPU", "SystemCPU","Elapsed", "Timelimit", "ReqMem", "CPUTimeRAW", "ElapsedRaw", "TimelimitRaw",
                  "TotalCPUSec", "CPUTimeSec", "ReqMemG", "JobID")

colnames.batch <- c("JobIDRaw", "Submit", "AllocNodes",
                    "AllocCPUS", "AveCPU", "MaxRSS", "MaxVMSize", "MaxDiskWrite", "MaxDiskWriteG",
                    "MaxDiskRead", "MaxDiskReadG", "MaxRSSG", "MaxVMSizeG", "JobID", "SubJob")

# Split out Job and Batch
batch.dt <- data[grepl("batch", SubJob), mget(colnames.batch)]
colnames(batch.dt)[colnames(batch.dt) %in% colnames.both] <-
  paste(colnames(batch.dt)[colnames(batch.dt) %in% colnames.both], "batch", sep = ".")

job.dt <- data[is.na(SubJob), mget(colnames.job)]
colnames(job.dt)[colnames(job.dt) %in% colnames.both] <-
  paste(colnames(job.dt)[colnames(job.dt) %in% colnames.both], "job", sep = ".")

# merge batch & job
cpu_only <- merge(x = job.dt, y = batch.dt, by = "JobID",
                  all.x = TRUE, all.y = FALSE, suffixes = c(".job", ".batch"))
# Calculate Efficiencies + Delay
cpu_only[, ':=' (CPUEff = round(TotalCPUSec/CPUTimeSec*100, 2),
                 MemEff = round(MaxVMSizeG/ReqMemG*100, 2), 
                 TimeEff = round(ElapsedRaw/(TimelimitRaw * 60) * 100, 2), 
                 Start = str_replace(Start, "T", " "), 
                 End = str_replace(End, "T", " "))]

cpu_only[, ':=' (Start = as.POSIXct(Start, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"), 
                 End = as.POSIXct(End, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"))]

cpu_only$Delay <- as.numeric(cpu_only$Start - cpu_only$Submit.job, units = 'secs')

cpu_only <- cpu_only %>% 
  mutate(CPUTimeHr =  CPUTimeSec / 3600) %>% 
  mutate(CPUCompUnits = CPUTimeHr / 6000) %>%  
  mutate(CPUCost = CPUCompUnits * 150)

cpu_only <- cpu_only %>% 
  mutate(State = if_else(str_detect(State, 'CANCELLED'), 'CANCELLED', State))

gpu_only <- cpu_only %>% 
  filter(Partition == 'gpu') %>% 
  mutate(ReqGPUS = as.numeric(str_extract(AllocTRES, "(?<=gpu=).*?(?=,|$)"))) %>%  
  mutate(GPUTimeSec = CPUTimeSec/ReqCPUS * ReqGPUS) %>%  
  mutate(GPUTimeHr = GPUTimeSec/3600) %>% 
  mutate(GPUCompUnits = GPUTimeHr / 600) %>%  
  mutate(GPUCost = GPUCompUnits * 150) 

total.dt <- merge(x = cpu_only, y = gpu_only %>%  select(JobID, ReqGPUS, GPUTimeSec, GPUTimeHr, GPUCompUnits, GPUCost) , by = "JobID", all.x = TRUE, all.y = FALSE)



```

## Summarize Data {.tabset}

### Summary by State and User
```{r}
print(paste0('Total # of Jobs: ', nrow(total.dt)))

# summarize by state 

DT::datatable(total.dt %>% group_by(State) %>% summarise(NumJobs = n(), CPUCost = round(sum(CPUCost, na.rm = TRUE),2), GPUCost = round(sum(GPUCost, na.rm = TRUE),2)), extensions = c('Buttons'), options = list(initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}"), searchHighlight = TRUE, scrollX = TRUE, scrollCollapse = TRUE))
```

```{r}
# SUmmarize by State by User 

filter <- c('State', 'User') 

if (length(unique(total.dt$User)) > 1) { 
  DT::datatable(total.dt %>% group_by(State, User) %>% summarise(NumJobs = n(), CPUCost = round(sum(CPUCost, na.rm = TRUE),2), GPUCost = round(sum(GPUCost, na.rm = TRUE), 2)), extensions =  c('Buttons'), options = list(initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}"), columnDefs = list(list(targets = filter, searchable = TRUE)), searchHighlight = TRUE, scrollX = TRUE, scrollCollapse = TRUE))
  }
```


### Summary by JobName 
Table displaying by User + JobName, the TimeEff, CPUEff and MemEff for each. 
```{r}
by_jobname <- total.dt %>%  
  group_by(JobName, User) %>% 
  summarise(TotalCPUSec = sum(TotalCPUSec, na.rm = TRUE), CPUTimeSec = sum(CPUTimeSec, na.rm = TRUE), ElapsedRaw = sum(ElapsedRaw, na.rm = TRUE), TimelimitRaw = sum(TimelimitRaw, na.rm = TRUE), MaxVMSizeG = sum(MaxVMSizeG, na.rm = TRUE), ReqMemG = sum(ReqMemG, na.rm = TRUE), NumJobs = n(), NumCompletedJobs = sum(State == 'COMPLETED'), CPUCost = round(sum(CPUCost, na.rm = TRUE),2) , GPUCost = round(sum(GPUCost, na.rm = TRUE),2)) %>%  
  mutate(TimeEff = ElapsedRaw/(TimelimitRaw * 60) * 100 ) %>%  
  mutate(CPUEff = TotalCPUSec/CPUTimeSec * 100) %>%  
  mutate(MemEff = MaxVMSizeG/ReqMemG*100) %>%  
  mutate(across(is.numeric, round, digits = 2))
  
  
DT::datatable( by_jobname %>% select(JobName, User, NumJobs, NumCompletedJobs, TimeEff, CPUEff, MemEff,  CPUCost, GPUCost), extensions =  c('Buttons'), options = list(initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}"), searchHighlight = TRUE, scrollX = TRUE, scrollCollapse = TRUE))

```


#### Scatter Plots by User for JobNames {.tabset}
Plots are only made for JobNames with > 5 Completed Jobs. 

To improve CPUEff, MemEff and TimeEff, decrease the requested # of CPUs, amount of memory and the timelimit accordingly! 

* Use horizontal scrollbar to view graphs fully! 

<style>
  .superbigimage{
      overflow-y:auto;
      overflow-x:auto;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
     max-height: none;
  }


</style>

NOTE: No plots will be shown if you specified '-all FALSE' in your command! 
```{r, results = 'asis', fig.width = 10, fig.height = 8}
if (params$all == 'TRUE'){ 

  for(x in unique((by_jobname %>% filter(NumCompletedJobs > 5))$User)) {
  
    cat('\n')
  
    cat('#####', x, '{.tabset}   \n')
  
    cat('\n')
  
    target_jobs <- unique((by_jobname %>% filter(NumCompletedJobs > 5 & User == x))$JobName)  
    
    user_info <- total.dt %>% filter(User == x & JobName %in% target_jobs) 
  
    subtabs <- c('CPUEff', 'MemEff', 'TimeEff')
    
    
    for(subtab in subtabs) {
      cat('\n')
  
      cat('###### ', subtab, '   \n')
  
      cat('\n')
      
        # if jobnames too many, need to split..
        chunks <- split(unique(user_info$JobName), ceiling(seq_along(unique(user_info$JobName)) / 8))
        
        for (x in chunks){ 
         subinfo <- user_info %>% filter(JobName %in% x)
         cat('<div class="superbigimage">')
          if (str_detect(subtab, 'CPU')){
           print(ggplot(data = subinfo, aes(x = ReqCPUS, y = CPUEff, colour = str_wrap(JobName, 10))) +
              geom_point()+ 
              theme_minimal() + 
              theme(panel.background = element_blank(), 
                    legend.title = element_blank(), 
                    plot.title = element_text(size = 15, face = 'bold'), 
                    axis.line = element_line(colour = 'black'), 
                    axis.text = element_text(size = 15.5), 
                    axis.title = element_text(size = 13, face = 'bold')) + 
              xlab("Requested CPUs") + ylab("CPU Eff (%)") + 
              scale_x_continuous(breaks = function(x) pretty(x, n = 8) %>% floor(), 
                                 )  + 
              coord_cartesian(clip = 'off') + 
               scale_color_brewer(palette="Dark2") + 
              ggtitle(paste0("User: ", unique(subinfo$User))))
            
          }
         
         else if (str_detect(subtab, 'Mem')){
        
       
          print(ggplot(data = subinfo, aes(x = ReqMemG, y = MemEff, colour = str_wrap(JobName, 10))) +
                  geom_point()+ 
                  theme_minimal() + 
                  theme(panel.background = element_blank(), 
                        legend.title = element_blank(), 
                        plot.title = element_text(size = 15, face = 'bold'), 
                        axis.line = element_line(colour = 'black'), 
                        axis.text = element_text(size = 15.5), 
                        axis.title = element_text(size = 13, face = 'bold')) + 
                  scale_x_continuous(breaks = function(x) pretty(x, n = 8) %>% floor()) +
                  xlab("Requested Memory (GB)") + ylab("Mem Eff (%)")  + 
                  scale_color_brewer(palette="Dark2") + 
                  ggtitle(paste0("User: ", unique(subinfo$User))) + 
                  coord_cartesian(clip = 'off'))
            
          
          }
            
      
      else {
        
         
        print(ggplot(data = subinfo, aes(x = TimelimitRaw, y = TimeEff, colour = str_wrap(JobName, 10))) +
              geom_point()+ 
              theme_minimal() + 
              theme(panel.background = element_blank(), 
                    legend.title = element_blank(), 
                    plot.title = element_text(size = 15, face = 'bold'), 
                    axis.line = element_line(colour = 'black'), 
                    axis.text = element_text(size = 15.5), 
                    axis.title = element_text(size = 13, face = 'bold')) + 
                xlab("Requested Timelimit (Mins)") + ylab("TimeEff (%)")  + 
              scale_x_continuous(breaks = function(x) pretty(x, n = 8) %>% floor()) +
               scale_color_brewer(palette="Dark2") + 
              ggtitle(paste0("User: ", unique(subinfo$User))) + 
              coord_cartesian(clip = 'off'))
    
         
      }
      
      cat('<div>')
      cat('\n')
      
        }
  
    }
  
    
  }

} 




```

### Full Table 
All sacct information and usage and cost calculations can be accessed here. 
Use Column Visibility to control what columns you want to see. 

*Refer to <a href = "https://slurm.schedmd.com/sacct.html">sacct slurm documentation</a> and <a href = 'https://ohsu-cedar-comp-hub.github.io/2023/09/19/Right-Sizing-Slurm-Jobs.html'>this Right Sizing page</a> for clarification. 

NOTE: No full table will be shown if you specified '-all FALSE' in your command! 

```{r warnings = FALSE}
if (params$all == 'TRUE'){ 
  # Reordering for ease 
  total.dt <- total.dt %>%  
    select(User, JobName, AllocTRES, CPUEff, MemEff, TimeEff, JobID, Account, Group, State, ReqCPUS, Submit.job, Submit.batch, Delay, Start, End, Elapsed, Timelimit, Cluster, Partition, SystemCPU, UserCPU, TotalCPU, TotalCPUSec, CPUTime, CPUTimeSec, MaxRSSG, MaxVMSizeG, MaxDiskReadG, MaxDiskWriteG,CPUTimeHr, CPUCompUnits, CPUCost, GPUTimeHr, GPUCompUnits, GPUCost) %>%  
    mutate(across(is.numeric, round, digits = 2))
  
  default_cols <- c('User', 'JobName', 'AllocTRES', 'CPUEff', 'MemEff', 'TimeEff', 'CPUCost', 'GPUCost')
  
  DT::datatable(total.dt, extensions = c('Buttons'), options = list(initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}"), searchHighlight = TRUE, 
      dom = 'Bfrtip',
        buttons = list('copy', 'print', list(
          extend = 'collection',
          buttons = c('csv', 'excel', 'pdf'),
          text = 'Download'
        ), I('colvis')), pageLength = 20, scrollX = TRUE, scrollCollapse = TRUE, 
        columnDefs = list(list(visible = FALSE, targets = colnames(total.dt)[!(colnames(total.dt) %in% default_cols)]))))     

  
} 
```
